CdGF_aucrf

### 2016-11-15 code/CdGF_AUCRF.Rmd
one mouse model leaving one source out

```{r}
pack_used <- c('randomForest','ggplot2', 'pROC', 'knitr','dplyr','AUCRF', 'tidyr')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

@@

meta_file   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'
shared_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared'
tax_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/mothur/gf_all.an.cons.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

@@

@@

@@

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='black', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
plot(LOdO_roc, col='red', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('Leave-2 cages-out (AUC: %.3g)',LOdO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc) #,
#                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
#                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,2,2,1,0,0), lwd=2, col=c('black','blue','purple','red','gray'), bty='n')
```

@@

```{r rel abund partial depend, fig.width=10, fig.height=10}

#n_rf_features <- 9
#euth_otus <- euth_otu_feat[1:n_rf_features]
euth_tax <- get_tax(1, euth_otu_feat, tax_file)
euth_tax$tax_label <- paste(euth_tax$tax, ' (', 
                            gsub('tu0*', 'TU ', rownames(euth_tax)), 
                                   ') : ', round(rf_early_aucrf$ranking[euth_otu_feat], 3), sep = '')

layout(matrix(c(1:9), ncol=3, byrow=TRUE))

for (i in seq_along(euth_otu_feat)) {
  partial_otu <- partialPlot(rf_early_aucrf$RFopt, aucrf_data, euth_otu_feat[i], xlab='',
                main=paste(euth_tax$tax_label[i]))
points(Predict_early_euth_df[Predict_early_euth_df$Euth_Early == 1 ,euth_otu_feat[i]], 
       jitter(rep(0.25*(diff(range(partial_otu$y))) + min(partial_otu$y), 
                  sum(Predict_early_euth_df$Euth_Early == 1)), 
              amount = (0.1*(diff(range(partial_otu$y))))), 
              pch = 21, bg = 'red')
points(Predict_early_euth_df[Predict_early_euth_df$Euth_Early == 0 ,euth_otu_feat[i]], 
       jitter(rep(0.75*(diff(range(partial_otu$y))) + min(partial_otu$y), 
                  sum(Predict_early_euth_df$Euth_Early == 0)), 
              amount = (0.1*(diff(range(partial_otu$y))))), 
              pch = 21, bg = 'gray')
  axis(side = 4, labels = c('Severe', 'Mild'),
       at = c(0.25*(diff(range(partial_otu$y))) + min(partial_otu$y),
              0.75*(diff(range(partial_otu$y))) + min(partial_otu$y)))
}
plot(NULL, xlim = c(0,1), ylim = c(0,1), xlab = '', ylab = '', axes = F)
legend('topleft', c('Mild', 'Severe'), pch = 21, pt.bg = c('gray','red'), cex = 2)
```

>>>>

```{r}
pack_used <- c('randomForest','ggplot2', 'pROC', 'knitr','dplyr','AUCRF', 'tidyr', 'caret')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

@@

meta_file   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'
shared_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared'
tax_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/gf_new.an.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

@@

# remove correlative otus
rel_abund_cor <- cor(rel_abund_stool, method = 'spearman')
cor_cutoff <- 0.75
cor_otus <- findCorrelation(rel_abund_cor, cutoff = cor_cutoff)
rel_abund_stool <- select(rel_abund_stool, -cor_otus)

@@

#### One_mouse model
Predict_om_df <- select(Predict_df, 
                        Euth_Early, human_source, mouse_id, contains('Otu'))
om_import_otus <- data.frame(otu = colnames(select(Predict_om_df, contains("Otu"))),
                             importance = NA)
om_aucs <- c()
om_all <- c()
for(j in 1:iters){
  set.seed(j)
  mouse_set <- Predict_om_df %>% 
    group_by(human_source) %>% 
    sample_n(1) %>% ungroup() 
  om_probs <- c()
  om_cage_import <- c()
  for(n in unique(as.character(mouse_set$human_source))){
    train <- mouse_set[mouse_set$human_source != n, ]
    test <- Predict_om_df %>% 
      filter(human_source %in% n)
    set.seed(seed)
    temp_model <- AUCRF(Euth_Early~., data=select(train, -mouse_id, -human_source)
                      , pdel=0.2, ntree=500, ranking = 'MDA')
    om_probs <- rbind(om_probs,
                      data.frame(obs = test$Euth_Early,
                                 pred = predict(temp_model$RFopt, test, type='prob')[,2]))
    om_cage_import <- c(om_cage_import, temp_model$Xopt)
  }
  om_roc <- roc(om_probs$obs~om_probs$pred)
  om_import_otus <- merge(om_import_otus,
                          data.frame(otu = names(table(om_cage_import)),
                            importance = as.numeric(table(om_cage_import))*om_roc$auc),
                          by = 'otu', all.x = T, suffixes = c('',round(om_roc$auc, 3)))
  om_all <- rbind(om_all, om_probs)
  om_aucs[j] <- om_roc$auc
}
om_roc <- roc(om_all$obs~om_all$pred)

raw_importance <- data.frame(om_import_otus[ , -c(1,2)], row.names = om_import_otus$otu)

importance_otus <-   apply(raw_importance, 1, sum, na.rm = T)
importance_otus <- importance_otus[importance_otus>0]
importance_otus <- importance_otus[importance_otus > median(om_aucs)*150/2]

### rank importance
importance_rank <- apply(raw_importance, 2, function(x) rank(x, na.last = T))
raw_rank <- apply(importance_rank, 1, function(x) sum(x, na.rm = T))
final_rank <- sort(raw_rank[raw_rank >=  ncol(importance_rank)*sum(importance_rank[ , 1])/nrow(importance_rank)], decreasing = T)

@@

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(om_roc, col='red', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
plot(LOdO_roc, col='red', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('Leave-2 cages-out (AUC: %.3g)',LOdO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc) #,
#                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
#                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,2,2,1,0,0), lwd=2, col=c('black','blue','purple','red','gray'), bty='n')
```

@@

```{r rel abund partial depend, fig.width=10, fig.height=10}
layout(matrix(c(1:9), ncol=3, byrow=TRUE))

for (i in seq_along(euth_otu_feat)) {
  partial_otu <- partialPlot(rf_early_aucrf$RFopt, aucrf_data, euth_otu_feat[i], xlab='',
                main=paste(get_tax(row_list = euth_otu_feat)$tax_label[i]))
points(Predict_early_euth_df[Predict_early_euth_df$Euth_Early == 1 ,euth_otu_feat[i]], 
       jitter(rep(0.25*(diff(range(partial_otu$y))) + min(partial_otu$y), 
                  sum(Predict_early_euth_df$Euth_Early == 1)), 
              amount = (0.1*(diff(range(partial_otu$y))))), 
              pch = 21, bg = 'red')
points(Predict_early_euth_df[Predict_early_euth_df$Euth_Early == 0 ,euth_otu_feat[i]], 
       jitter(rep(0.75*(diff(range(partial_otu$y))) + min(partial_otu$y), 
                  sum(Predict_early_euth_df$Euth_Early == 0)), 
              amount = (0.1*(diff(range(partial_otu$y))))), 
              pch = 21, bg = 'gray')
  axis(side = 4, labels = c('Severe', 'Mild'),
       at = c(0.25*(diff(range(partial_otu$y))) + min(partial_otu$y),
              0.75*(diff(range(partial_otu$y))) + min(partial_otu$y)))
}
plot(NULL, xlim = c(0,1), ylim = c(0,1), xlab = '', ylab = '', axes = F)
legend('topleft', c('Mild', 'Severe'), pch = 21, pt.bg = c('gray','red'), cex = 2)
```
